Synthesising Recursive Functions for\\ First-Order Model Counting

### 2023-02-26 doc/talks/3_long/talk.tex
Some writing and some more feedback

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{complexity}
\usepackage[style=authoryear]{biblatex}
\usepackage{forest}
\usepackage{soul}
\usepackage{csquotes}
\usepackage{pifont}

@@

\definecolor{color1}{HTML}{1B9E77}
\definecolor{color2}{HTML}{7570b3}
\definecolor{color3}{HTML}{e7298a}
\definecolor{forestgreen}{HTML}{009B55}

@@

\newcommand{\cmark}{\ding{51}}%
\newcommand{\xx}{\textcolor{color1}{x}}
\newcommand{\yy}{\textcolor{color3}{y}}
\newcommand{\pP}{\texttt{P}}
\newcommand{\qQ}{\texttt{Q}}
\newcommand{\dDelta}{\textcolor{color2}{\Delta}}

@@

\author{\textbf{Paulius Dilkas} \inst{1} \and Vaishak Belle \inst{2}}
\title{Recursive Solutions to \\First-Order Model Counting}
\institute{\inst{1} National University of Singapore, Singapore \and \inst{2} University of Edinburgh, UK}
\date{TODO: date}

@@

% TODO: add a very clear example of what a model is
% TODO: introduce the problem from the next side, visualise P and Q as subsets of Delta?
% TODO: (maybe) add a slide about other approaches to WFOMC, making it crystal clear what my contributions are
\begin{frame}{First-Order Model Counting (in More Detail)}
  \begin{itemize}
    \item \alert{First-order model counting} (FOMC) is the problem of counting
          the models of a sentence in first-order logic.
    \item The \alert{(symmetric) weighted} variation of the problem adds weights
          (e.g., probabilities) to predicates.
          \begin{itemize}
            \item It is used for efficient \alert{probabilistic inference} in
                  relational models such as Markov logic networks.
          \end{itemize}
    \item None of the (implemented) (W)FOMC algorithms are able to count, e.g.,
          \alert{injective} and \alert{bijective} functions.
  \end{itemize}
  \begin{block}{Claim}
    This shortcoming can be addressed via support for \alert{recursive functions}.
  \end{block}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{complexity}
\usepackage[style=authoryear]{biblatex}
\usepackage{forest}
\usepackage{soul}
\usepackage{csquotes}
\usepackage{pifont}
\usepackage{venndiagram}

@@

\definecolor{color1}{HTML}{1B9E77}
\definecolor{color2}{HTML}{7570b3}
\definecolor{color3}{HTML}{e7298a} % TODO: I don't like pink
\definecolor{forestgreen}{HTML}{009B55}

@@

\newcommand{\cmark}{\ding{51}}%

\newcommand{\xx}{\textcolor{color1}{x}}
\newcommand{\yy}{\textcolor{color3}{y}}
\newcommand{\pP}{\texttt{P}}
\newcommand{\qQ}{\texttt{Q}}
\newcommand{\dDelta}{\textcolor{color2}{\Delta}}

\newcommand{\pone}{\textcolor{color1}{\pP(1)}}
\newcommand{\ptwo}{\textcolor{color1}{\pP(2)}}
\newcommand{\qone}{\textcolor{color2}{\qQ(1)}}
\newcommand{\qtwo}{\textcolor{color2}{\qQ(2)}}
\newcommand{\pqone}{\textcolor{color3}{\pP(1)}, \textcolor{color3}{\qQ(1)}}
\newcommand{\pqtwo}{\textcolor{color3}{\pP(2)}, \textcolor{color3}{\qQ(2)}}

@@

\author{\textbf{Paulius Dilkas} \inst{1} \and Vaishak Belle \inst{2}}
\title{Synthesising Recursive Functions for\\ First-Order Model Counting}
\institute{\inst{1} National University of Singapore, Singapore \and \inst{2} University of Edinburgh, UK}
\date{TODO: date}

@@

% TODO: (maybe) add a slide about other approaches to WFOMC, making it crystal clear what my contributions are

\begin{frame}{(Unweighted) First-Order Model Counting}
  \begin{columns}[T]
    \begin{column}{0.5\textwidth}
      \begin{itemize}
        \item Example formula: $\forall x \in \Delta\text{. } \pP(x) \lor \qQ(x)$
        \item Let $\Delta \coloneqq \{\, 1, 2 \,\}$
        \item Interpretations: all subsets of $\{\, \pP(1), \qQ(1), \pP(2), \qQ(2) \,\}$
        \item Models:
        \[
          \begin{matrix}
            \{\, \pone, \ptwo \,\}, & \{\, \pone, \qtwo \,\}, & \{\, \pone, \pqtwo \,\},\\
            \{\, \qone, \ptwo \,\}, & \{\, \qone, \qtwo \,\}, & \{\, \qone, \pqtwo \,\},\\
            \{\, \pqone, \ptwo \,\}, & \{\, \pqone, \qtwo \,\}, & \{\, \pqone, \pqtwo \,\}\\
          \end{matrix}
        \]
        \item Each 1-ary predicate...
      \end{itemize}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{overlayarea}{\textwidth}{4cm}
        \only<1>{
          \begin{venndiagram2sets}[labelNotAB={$\Delta$},labelA={$\pP$},labelB={$\qQ$}]
          \end{venndiagram2sets}
        }%
        \only<2>{
          \begin{venndiagram2sets}[labelNotAB={$\Delta$},labelA={$\pP$},labelB={$\qQ$}]
            \fillOnlyA;
          \end{venndiagram2sets}
        }%
        \only<3>{
          \begin{venndiagram2sets}[labelNotAB={$\Delta$},labelA={$\pP$},labelB={$\qQ$}]
            \fillOnlyB;
          \end{venndiagram2sets}
        }%
        \only<4>{
          \begin{venndiagram2sets}[labelNotAB={$\Delta$},labelA={$\pP$},labelB={$\qQ$}]
            \fillACapB;
          \end{venndiagram2sets}
        }%
        \only<5>{
          \begin{venndiagram2sets}[labelNotAB={$\Delta$},labelA={$\pP$},labelB={$\qQ$}]
            \fillA;
            \fillB;
          \end{venndiagram2sets}
        }%
      \end{overlayarea}
    \end{column}
  \end{columns}
\end{frame}
% TODO: mention that n-ary predicates ~= relations

% \begin{frame}{First-Order Model Counting (in More Detail)}
%   \begin{itemize}
%     \item \alert{First-order model counting} (FOMC) is the problem of counting
%           the models of a sentence in first-order logic.
%     \item The \alert{(symmetric) weighted} variation of the problem adds weights
%           (e.g., probabilities) to predicates.
%           \begin{itemize}
%             \item It is used for efficient \alert{probabilistic inference} in
%                   relational models such as Markov logic networks.
%           \end{itemize}
%     \item None of the (implemented) (W)FOMC algorithms are able to count, e.g.,
%           \alert{injective} and \alert{bijective} functions.
%   \end{itemize}
%   \begin{block}{Claim}
%     This shortcoming can be addressed via support for \alert{recursive functions}.
%   \end{block}
% \end{frame}
